---
title: "Homework 2"
author: "simplymathematics"
date: "September 22, 2018"
output: html_document
---
Graded: 2.6, 2.8, 2.20, 2.30, 2.38, 2.44


## 2.6
- A pair of dice will never give you a one.
- Since there are 36 ways to roll 2 dice and 4 ways to rolls a 5 (1 + 4, 4+1, 2+3, 3+2), there is a 5/36 chance of rolling a 5.
- Since there are 36 ways to roll 2 dice and only 1 way to get a 12 (6+6), the probability is 1/36. 

## 2.8
- Poverty and speaking an language other than English are not disjoint
- Draw a venn diagram: 14.6 in poverty, 20.7, and 4.2 both

```{r}
poverty <- 14.6
ESL <- 20.7
both <- 4.2
library(VennDiagram)
venn.plot <- draw.pairwise.venn(14.6, 20.7, 4.2, c("Poverty", "Not English"))
grid.draw(venn.plot)
grid.newpage();

```
- 10.4%  live in poverty and speak only English at home 
- 16.5+4.2+10.4 = 31.1% live in poverty or speak a foreign language at home
- 85.4% don't live in poverty and 79.3% only speak English at home. Because these are independent events, their total probability is multiplied and = 69%
- Independence relies on any one of these equivalent conditions
$$
P(A \cap B) = P(A|B)P(B)  \\
P(A|B) = P(A)             \\
P(B|A) = P(B)
$$ 
if and only if A and B are independent. If we assume that A is the ESL status and B is the poverty level, then we see that 

```{r}
p_a = 20.7
p_b = 14.6
p_a_and_b = 4.2
p_a_given_b = p_a_and_b/p_b
p_a_given_b
p_a
```

$$
P(B|A) \neq P(B)
$$

## 2.20
```{r}

df <- data.frame(matrix(c(78,19,11,108, 23,23,9,55,13,12,16,41,114,54,36,204),nrow=4,ncol=4))

rownames(df) <- c("mblue", "mbrown", "mgreen", "totals")
colnames(df) <- c("fblue", "fbrown", "fgreen", "totals")

df
  
```

- Probability that male or partner has blue eyes is 71%. Where m is a male, f is a female, b represents blue eyes, and br represents brown. 
```{r}
mb = 114
fb = 108
bfm = 78
answer = (mb+ fb-bfm)/204
answer
```

- The probability that the partner has blue eyes given the male has blue eyes is .68.
```{r}
p_a = fb
p_b = mb
p_a_and_b = bfm
p_a_given_b = p_a_and_b/p_b
p_a_given_b
```
- They do not appear to be independent because the below is
```{r}
p_a == p_a_given_b
```
- Probability of fb | mbr .35
```{r}
pfb = 108/204
pmbr = 54/204
fbmbr = 19/204
a = pfb
b = pmbr
ab = fbmbr
pagb = ab/b
pagb

```
- Independence? No
```{r}
a == pagb
```
## 2.30
```{r}
df <- data.frame(matrix(c(13,15,28,59,8,67,72,23,95),nrow=3,ncol=3))
rownames(df) <- c("fiction", "non-fiction", "totals")
colnames(df) <- c("hardcover", "paperback", "totals")
df
  
```
- Probability of drawing a hardcover book, then a paperback fiction without replacement .21
```{r}
(28/95)*(67/94)
```

- The probabilty of fiction book 1st then hardcover 2nd, without replacement, is .2243:
```{r}
fiction_and_hardcover = 13/95
fiction_not_hardcover = 59/95
second_fiction_given_fh = 27/94
second_fiction_given_nfh = 28/94
h_after_fh = second_fiction_given_fh*fiction_and_hardcover
h_after_nfh  =second_fiction_given_nfh*fiction_not_hardcover
answer = h_after_nfh+h_after_fh
answer
```
- With replacement, it is .2234
```{r}
(72/95)*(28/95)
```
- The sample is large enough that a single book doesn't much effect the results.

## 2.38
We know that revenues are \$25 for first bag and \$35 or the 2nd (making revenue \$60 for each person who checks 2 bags). We also know that 54% people check no bags, 34% check 1, and 12% check 2. The expected per passenger is
```{r}
df <- data.frame(matrix(c(0,25,60,.54,.34,.12),nrow=3,ncol=2))
colnames(df) <- c("x","P(x)")
df
```
The expected value is
```{r}
expected = 0*.54 + 25*.34+.12*60
expected
```
Likewise, the standard deviation can be found by
```{r}
prices = c(0,25,60)
means = c(.54,.34,.12)
vec = c(prices - expected)
squares = c(vec * vec)
squares %*% means
(squares %*% means)^.5
```

## 2.44

```{r}
df <- data.frame(matrix(c(9999, 14999, 24999, 34999, 49999, 64999, 74999, 99999, 100000,.022,.047,.158,.183,.212,.139,.058,.084,.097),nrow=9,ncol=2))
colnames(df) <- c("income up to","total")
df
barplot(df$total)
```

Personal income tends to be grouped near \$43k while a sizeable amount of people also make more than \$75k. In fact, 62% of people make less than $50,000. 
```{r}
sum(df$`total`[0:5])
```
If we assume that gender is independent of income, then the probability of a woman making less than $50,000 is the same as the probability for the total population. However, that population is only 1/2 female. So, we'd multiply these two seemingly independent events together to get
```{r}
.622 * .5
```
However since we know that the probability of a person being a woman given that they make <\$50k a year is 71.8%, we know that these events are not independent becase
$$ P(A) \neq P(A|B) $$ s


