---
title: "606 Chapter 1"
author: "simplymathematics"
date: "September 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 1 Homework

## Setup

```{r}
# Uncomment this line if you don't have the openintro package
# install.packages('openintro')
library(openintro); data(heartTr)
head(heartTr,100)
```
## 1.8
### A
Each row represents a person in the dataset
### B
```{r}
nrow(datas)
```
### C

id: numerical, discrete
sex: categorical
marital: categorical
grossIncome: categorical, ordinal
smoke: categorical
amtWeekends: numerical, discrete
amntWeekdays: numerical, discrete

## 1.10

The population is children between 5 and 15. The sample is 160 of them. However, this study would be poorly generalizable because there is no randomization of the test/control groups, there is no randomized sample selection, and we have no idea how these children were selected. Imagine how the results in this sample might skew a study done on students at an accelerated school versus the general population.
## 1.48
```{r}
nums = c(57,66,69,71,72,73,74,77,78,78,79,79,81,81,82,83,83,88,89,94)
summary(nums)
boxplot(nums, main= "Scores of 20 statistics students", ylab = "Scores (%)")
```
## 1.50
A is basically normally distributed, which corresponds to 2. B is pretty uniformly distributed, which follows box plot 3. C is distributed like a Poisson distribution, which corresponds to boxplot 1. 
## 1.56

a. skewed right, IQR and median
b. symmetric, mean and std
c. skewed left, IQR and median
d. skewed left, IQR and median
## 1.70

### A
The mosaic shows that a person receiving treatment was much more like to survive in the long-term.
### B
The median and mean survival times for a person who received a transplant are much higher than those who didn't. However, the treatment is not always successful. On average, a transplant adda bout 200 days of survival time.
### C
Approximately 4/5 of the control group patients died and only 2/3 of the transplant patients died over the course of the study.
### D
i. The claim being tested is "Heart transplants add survival time for these patients"
ii. 
```{r}
#nrow(heartTr)
nrow(subset(heartTr, survived =='dead'))
nrow(subset(heartTr, survived =='alive'))
a = nrow(subset(heartTr, transplant == 'treatment'))
a
b = nrow(subset(heartTr, transplant =='control'))
b
a/b
# ...were the simulated differents in proportions are different
```
iii. The simulations below show that the transplant program is likely effective.