---
title: "Chapter 6 606"
author: "simplymathematics"
date: "November 4, 2018"
output: html_document
---

# 6.6
On June 28, 2012 the U.S. Supreme Court upheld the much debated 2010 healthcare law, declaring it constitutional. A Gallup poll released the day after this decision
indicates that 46% of 1,012 Americans agree with this decision. At a 95% confidence level, this sample has a 3% margin of error. Based on this information, determine if the following statements are true or false, and explain your reasoning.

False
We are 100% certain that the percentage  of Americans in this sample support that support the decision is 46%.


True
We are 95% confident that between 43% and 49% of Americans support the decision of the
U.S. Supreme Court on the 2010 healthcare law because the 95% CI is [43%,49%].

False
If we considered random samples, we could say nothing of the confidence interval of the sample data using the population data.

False
If the confidence level decreases, then the margin of error would also decrease. 

# 6.12
```{r]}
p.hat = .48
n = 1259
``` 
A parameter is a descriptive measure for a population-- a statistic describes a sample.

For $ C.I. = 95%$

$z* =1.96$

Then the margin of error is
$$
ME = z^*\cdot\sqrt{\frac{\hat p(1-\hat p)}{n}} = .0276
$$
p + ME = 51%
p - ME = 45%


The normal is a good approximation because np and n(1-p) are both greater than 10.

The CI contains the .50 and thus is appears that 50% of Americans think marijuana should be legalized. Since 50% is not more than half, the piece's statement wasn't justified.

# 6.20
```{r}
p.hat = .48
ci = .95
ME = .02
z = 1.96
n = (z*sqrt(p.hat*(1-p.hat))/ME)^2
ceiling(n)
```
We would need to survey about 2400 Americans.

# 6.28
```{r}
n1 = 11545
p1 = .08
n2 = 4691
p2 =.088
c = .95
z = 1.96

ME = z*sqrt((p1*(1-p1)/n1)+((p2*(1-p2)/n2)))
ME
bottom = p1 - p2 - ME
top = p1 -p2 + ME
interval = round(c(bottom, top),4)*100
interval
```
We are 95% confident that the true proportion of Californians who lack sleep is within this interval of percentage from the Oregonian mean.


# 6.44

```{r}
n = 426
a = .05
```
H0: The population proportions are as mentioned.  
Ha: The population proportions are different than mentioned.  

When we want to see how 1 variable varies with another, we use a $ \chi^2 $ test.  

We must assume the samples are independent of each other. 

$$
\chi^2 = \sum \frac{E}{(O-E)^2}=284 \\
df = c - 1 = 3
$$

```{r}
observed = c(4,16,61,345)/426
expected = c(.048,.147,.396, 1-.048-.147-.396)
chisq.test(x = observed, p = expected)
```
We can reject the null hypothesis.

# 6.48
## a
We will use a $\chi^2$ goodness of fit test with an $\alpha = .05$  
## b
H0: The variables are independent  
Ha: The variables are dependent  

## c
```{r}
ratio = 2607/48132
ratio
```
5% of women suffer from depression.
## d
```{r}
ratio * 6224
```

## e
```{r}
1-pchisq(20.93, df = 4)
```

## f
We can reject the null hypothesis because there is sufficient evidence that clinical depress is related to caffeinated coffee consumption.

## g
Yes. It appears that coffee may influence one's propensity for depression.