---
title: "606 Chapter 4"
author: "simplymathematics"
date: "October 21, 2018"
output: html_document
---
# Chapter 4

## 4.4
### a 
Mean : 171.1
Median : 170.3
### b
sd: 9.4
IQR: 177.8-163.8 = 14
### c
A 180 cm person would be among the tallest 25%. A person who is 155cm tall is among the shortest 25%. 
### d
We would expect them to be different because we would be sampling a different population.
### e
The variability is 9.4/(507)^.5  = .401

## 4.14
### a
True, because the the 95% confidence interval is between 80.41 and 89.11.
### b 
False. The sample skew does not invalidate a sample mean.
### c 
True, because that's what confidence interval means.
### d 
False. This this is not a survey of total spending, just holiday spending.
### e 
True. a 90% Confidence Interval would be smaller because it would cover fewer options.
### f
To decrease the margin of error by 2/3, we'd have to have 9 times the sample size. 

### g
The standard  error is +/-
$$
\frac{\sigma}{\sqrt{n}} = \frac{\sigma}{\sqrt{436}} \text{ where} \\
\sigma = \frac{\frac{89.11+80.42}{2}-80.42}{1.96} = 4.35 \\
\rightarrow \frac{\sigma}{\sqrt{n}} = .21
$$

## 4.24

### a
Yes. the study used random, independednt samples of a sufficient number (30 $n$ < 10% population).

### b
```{r}
mean = 30.69
sd = 4.31
xbar = 32
n = 36
s = sd/sqrt(n)
z <- (mean-xbar)/s
z
pnorm(-abs(z))
```
There is sufficient evidence to support the claim that gifted children read earlier. If the population mean is 32 months, we have a 3.4% of obtaining a sample of 36 children with an average of 30.69 months.
```{r}
mean = 30.69
sd = 4.31
s = sd/sqrt(n)
upper <- mean + s * 1.65
lower <- mean - s * 1.65
CI <- c(lower, upper)
CI
```
They agree because they describe the same data from two viewpoints.
## 4.26
```{r}
mean = 118.2
sd = 6.5
n = 36
s = sd/sqrt(n)
a = .1
z <- (xbar-100)/s
z
pnorm(z)
```
There is sufficient evidence to support the claim that the average IQ of mothers of gifted children is different than that of the general population.

```{r}

mean = 118.2
sd = 6.5
n = 36

s = sd/sqrt(n)
upper <- mean + s * 1.65
lower <- mean - s * 1.65
CI <- c(lower, upper)
CI
```

## 4.34
A "sampling distribution" is vector that contains on the sample means.

The CLT tells us that the sampling distribution approaches the normal distribution if the sample size is at least 30.

The mean sampling distribution is equal to the population mean.

The standard deviation of the sampling distribution of the sample mean is 
$$
\frac{\sigma}{\sqrt{n}}
$$
As the sample size increases, the shape of the sampling distribution becomes more normal.
## 4.40
```{r}
mean = 9000
sd = 1000
z = (10500 - mean)/sd
1 -pnorm(z)
```
The P(x = 10500) is 6%. 

```{r}
n = 15
s = sd/sqrt(n)
s
```
The standard deviation of the sampling distribution is 258.

```{r}
z = (10500-mean)/(s)
z
1- pnorm(z)
```
There is a miniscule probability associated with a z-score of 5.8.
## 4.48
As the sample size increases, the absolute value of Z increases. This decreases the probaility.
