---
title: "606 Final"
author: "simplymathematics"
date: "December 13, 2018"
output: html_document
---

# Final 

## Part 1

### Problem 1

daysDrive is quantitative and discrete since it is an integer that represents the number of days a week a student drives to school.

### Problem 2

The plot reveals that the data is skewed right, meaning the mean is higher than the median. This only leaves B and D as potential answers. However, we can see that the median is likely higher than 2.9, so we are left with B.

### Problem 3

A randomized, controlled trial is the correct way to determine the efficacy of a drug. This is option A. An observational study like is proposed in B would suffer from survivorship bias. People who don't survive Ebola would never make it into the statistics.

### Problem 4

If a large $X^2$ value is obtained between two vectors representing hair and eye color, then we can say there is a relationship between the two.

### Problem 5
Boxplots show the IQR or the range from Q1 to Q3. Therefore, in order to see if a score is an outlier we need those two numbers. In this cases it is 37 and 49.8 (A). 

### Problem 6

Median and IQR are resistant to outliers. Mean and standard deviation are not.

### Problem 7

#### a 

The observation distribution is  nearly normally distributed, but is skewed left.  Thesampling distribution is more symmetric, but has smaller standard deviation.

#### b

The means of these two disitributions are similar even though their standard deviations differ because standard deviation is calculated as the sum of residuals squared. Since the observation distribution has a smaller number of instances, we would expect its sum of residuals to be smaller. Likewise, we would expect a smaller standard deviation.

#### c
The central limit theorem says that as we add more observations of independent random variables, their normalized sum tends towards the normal distribution. 


## Part 2
```{r}
x1 <- c(10,8,13,9,11,14,6,4,12,7,5)
y1 <- c(8.04, 6.95, 7.58, 8.81, 8.33, 9.96, 7.24, 4.26, 10.84, 4.82, 5.68)

```

### a
```{r}
mean(x1)
mean(y1)
```

### b
```{r}
median(x1)
median(y1)
```
### c
```{r}
sd(x1)
sd(y1)
```
### d
```{r}
cor(x1,y1)
```
### e
```{r}
model1 <- lm(y1~x1)

model1
```
### f
```{r}
summary(model1)
```
### g
```{r}
plot(x1,y1)
abline(model1)
plot(model1)
```
### h